---
title: "Housing Data During the '07 Market Crash"
author: "Austin Huber"
date: "Due December 17, 2019"
output:  html_notebook
---

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(ggplot2)
library(mosaic)
library(mosaicData)
library(DataComputing)
library(tidyverse)
library(tidyr)
library(lubridate)
library(readr)
```


## Research Question:

How did the house prices of houses in America change during the '07 market crash? Are there specific attributes about houses that contributed to them being able to hold their value better through this time period? 


## Data Access

Loading the data as a whole into a new variable in the RStudio environment. Here is the link of where the data was acquired from: https://www.kaggle.com/gpandi007/usa-housing-dataset
```{r include=FALSE}
#Housingdata <- read_csv("housing_train.csv")
Saleprices <- read_csv("Sale_Prices_Msa.csv")
```
```{r}

nrow(Saleprices)
ncol(Saleprices)
names(Saleprices)
head(Saleprices)

```

This shows the average house price across the US 

```{r}

Avgdata <- Saleprices %>% filter(RegionID==102001)
newAvgdata <- Avgdata %>% gather(month, price, -RegionID, -RegionName, -SizeRank)

newAvgdata %>%
  mutate(Year = lubridate::dym(month)) %>%
  ggplot(aes(x=Year,y=price))+geom_line() 
```




