---
title: "Housing Data After the 2007-08 Market Crash"
author: "Austin Huber"
date: "December 10, 2019"
output:  html_notebook
---

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(ggplot2)
library(mosaic)
library(mosaicData)
library(DataComputing)
library(tidyverse)
library(tidyr)
library(lubridate)
library(readr)
```

```{r include=FALSE}
Saleprices <- read_csv("Sale_Prices_Msa.csv")
DaysUp <- read_csv("DaysOnZillow_State.csv")
```

## Research Question:

How long did the recovery of the housing market take in the US after the 2007-08 market crash? How have prices changed since then? Does location have any impact on the recovery of housing prices?


## Challenge Encountered:

How long did the recovery of the housing market take in the US after the 2007 market crash? How have prices changed since then? Could there be another bubble burst coming soon?


## Key Visualization:

```{r}
JustMonths <- Saleprices[-c(1,2,3)]
Mean <- apply(JustMonths,2,mean,na.rm=TRUE)
Max <- apply(JustMonths,2,max,na.rm=TRUE)
Min <- apply(JustMonths,2,min,na.rm=TRUE)
MonthlyCalcs <- data.frame("month"=colnames(Justmonths),Min,Max,Mean)
USdata <- Saleprices %>% filter(RegionID==102001) %>% gather(month, price, -RegionID, -RegionName, -SizeRank) %>% mutate(Month = lubridate::dym(month))
MonthlyCalcs %>%
  mutate(Month = lubridate::dym(month)) %>%
  ggplot(aes(x=Month,y=Mean,color="Average in top 626 Cities"))+geom_line(size=1)+geom_line(mapping=aes(y=price,color="Average in all of the US"),data=USdata,size=1)+ggtitle("Average Price of Houses in the US")+xlab("Years")+ylab("Price (in dollars)")
```

## Key Insight/Takeaway:

How long did the recovery of the housing market take in the US after the 2007 market crash? How have prices changed since then? Could there be another bubble burst coming soon?




## Data Access

Loading the data as a whole into a new variable in the RStudio environment. Here is the link of where the data was acquired from: https://www.zillow.com/research/data/
```{r include=FALSE}
Saleprices <- read_csv("Sale_Prices_Msa.csv")
DaysUp <- read_csv("DaysOnZillow_Msa.csv")
```
```{r}
cat("Number of rows: ",nrow(Saleprices),"\n")
cat("Number of columns: ",ncol(Saleprices),"\n")
head(Saleprices)
cat("Number of rows: ",nrow(DaysUp),"\n")
cat("Number of columns: ",ncol(DaysUp),"\n")
head(DaysUp)
```
