---
title: "Housing Data During the '07 Market Crash"
author: "Austin Huber"
date: "Due December 17, 2019"
output:  html_notebook
---

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(ggplot2)
library(mosaic)
library(mosaicData)
library(DataComputing)
library(tidyverse)
library(tidyr)
library(lubridate)
library(readr)
```


## Research Question:

How long did the recovery of the housing market take in the US after the 2007 market crash? How have prices changed since then? Could there be another bubble burst coming soon?


## Data Access

Loading the data as a whole into a new variable in the RStudio environment. Here is the link of where the data was acquired from: https://www.zillow.com/research/data/
```{r include=FALSE}
Saleprices <- read_csv("Sale_Prices_Msa.csv")
#NoHeader <- read_csv("Sale_Prices_Msa.csv",skip=1)
```
```{r}
cat("Number of rows: ",nrow(Saleprices),"\n")
cat("Number of columns: ",ncol(Saleprices),"\n")
head(Saleprices)
```

```{r}
JustMonths <- Saleprices[-c(1,2,3)]
MeanMonth <- apply(JustMonths,2,mean,na.rm=TRUE)
MaxMonth <- apply(JustMonths,2,max,na.rm=TRUE)
MinMonth <- apply(JustMonths,2,min,na.rm=TRUE)
TempArr <- data.frame("Month"=colnames(Justmonths),MinMonth,MaxMonth,MeanMonth)
MonthlyCalcs <- data.frame(t(TempArr[-1]))
colnames(MonthlyCalcs) <- TempArr[,1]
TotalInfo <- 
  Saleprices %>%
  mutate(MinimumPrice = apply(Justmonths,1,min,na.rm=TRUE)) %>%
  mutate(MaximumPrice = apply(Justmonths,1,max,na.rm=TRUE)) %>%
  mutate(MeanPrice = apply(Justmonths,1,mean,na.rm=TRUE))
TotalInfo
MonthlyCalcs
```


This shows the average house price across the US 

```{r}
USdata <- Saleprices %>% filter(RegionID==102001) %>% gather(month, price, -RegionID, -RegionName, -SizeRank)
USdata %>%
  mutate(Year = lubridate::dym(month)) %>%
  ggplot(aes(x=Year,y=price))+geom_line() 
Totalin
```




